\input{macros/preamble}

\begin{document}

\maketitle% this prints the handout title, author, and date

%\begin{fullwidth}
%\begin{multicols}{2}
\tableofcontents\marginnote{(Acknowledgements will go in a margin note here.)}
%\end{multicols}
%\end{fullwidth}

\newpage
\vfill
\begin{myboxB}
\centering
\textbf{Novel contributions:}
\begin{itemize}

\item Section \bR REF \e is a pedestrian introduction to weak $n$-category theory (via \texttt{homotopy.io}, underpinned by the theory of associative $n$-categories) from the perspective of generalising familiar string-rewrite systems to higher dimensions. The chief development of this section is a demonstration that context-free grammars and tree-adjoining grammars may be formalised in the $n$-categorical setting.

\item Section \bR REF \e spells out a generative grammar for text using an $n$-categorical signature as a rewrite system, which additionally provides a unified framework from which the Text Circuit Theorem first proved in \bR CITE \e is recovered.

\item Section \bR REF \e introduces the category \textbf{ContRel} of continuous relations. I detail the relationships (or lack thereof) of \textbf{ContRel} to its cousins \textbf{Top} and \textbf{Rel}. Though \textbf{ContRel} is constructed na\"{i}vely, its definition and an exposition of its expressivity from the monoidal perspective appears to be novel.

\item Section \bR REF \e string-diagrammatically characterises set-indexed collections of disjoint open subsets of spaces in \textbf{ContRel} as \emph{sticky spiders} -- special frobenius algebras that satisfy certain interaction relations with an idempotent. The diagrammatic outcome is that reasoning with such set-indexed collections remains as graphically intuitive as with spiders.

\item Section \bR REF \e -- with the aid of a theorem by Friedman \citep{friedman_fom_2005} -- string-diagrammatically characterises a vocabulary of linguistic topological relations in \textbf{ContRel} such as simple connectedness, touching, insideness, and rigid motion.

\item Section \bR REF \e argues for the centrality of explaining communication as a criterion for formal approaches to syntax, and explores the relationship between productive and parsing grammars as organised by a monoidal cofunctor. A diagrammatic treatment of monoidal cofunctor boxes is introduced for this purpose.

\item \bR REF \e is a standalone introduction to the mathematical setup of Montague's \emph{Universal Grammar}, aimed at modern algebraists. An outcome of this section is the placement of text circuits as a natural mathematical development in the broadly conceived programme of Montague Semantics.

\item Appendix \bR REF \e constructs a subcategory of sticky spiders in \textbf{ContRel} on the unit square that behaves as a model of \textbf{FinRel} equipped with a \emph{Turing object} -- a single object in a category with respect to which all other objects and morphisms between them may be encoded. This is of particular relevance to the linguistic phenomenon of \emph{entification} -- that essentially all grammatical categories of words and even phrases have noun-equivalents, e.g. $\texttt{to run} \simeq \texttt{running}$, $\texttt{Bob drinks} \simeq \texttt{the fact that Bob drinks}$. The presence of a Turing object in a symmetric monoidal category additionally provides a semantic model for higher-order processes of generic input type, and for \emph{lasso diagrams}.

\item In Section \bR REF \e, by parsing text as circuits (Section \bR REF \e) and using monoidal cofunctor boxes (Section \bR REF \e) to interpret those circuits in \textbf{ContRel} as iconic representations equipped with a vocabulary of lingustic-topological concepts (Section \bR REF \e), I compute some metaphors by hand.

\end{itemize}
\end{myboxB}
\vfill
\clearpage
\newpage

\setcounter{chapter}{-1}
\chapter{Context and synopsis}
There are potentially practical and theoretical benefits to a fresh mathematical take on basic linguistics. String diagrams are formal, intuitive, expressive, fun, and pretty. I review the relevant research context.
\clearpage
\newpage
\input{chapters/intro/intro_draft3}
\newpage
\input{chapters/stringdiagrams/1b_proctheory}\label{sec:proctheory}
\newpage
\input{chapters/textcirc/previously}



\bibliographystyle{alpha}
\bibliography{thesis_intro}





\end{document}