\section{Postscript: On meaning and meaninglessness}

The version of the thesis my examiners passed began with a remark to the effect that I share an interest with (certain) linguists in language as a (potentially) formal object, but I would not hold myself hostage to their literature, nor their methods. My examiners correctly deduced that this dismissive attitude was in part a cover to avoid engaging with the literature in linguistics altogether, which led to a brief sidebar on whether there are kinds of mathematicisation that qualify as intellectual colonialism. The consensus was yes, and on that note, I do share some methods in common with the linguists at MIT between \emph{Syntactic Structures} and \emph{Aspects of the Theory of Syntax}, as well as the stereotypical physicist encountering new fields with fresh eyes [XKCD] -- namely, armchair introspection and its sidekick, synoptic and reductive mathematical "toy-modelling" in search of simplicity, along with a healthy disregard for what others think. I gave two other reasons for the lack of engagement, but my arguments that I shouldn't be obliged to cite unreadable and useless material because I view myself as a kind of above-it-all self-loathing artist-cum-diagram-arms-dealer was dismissed by my examiners, who pointed out my obligations as a scholar. Now that I have satisfied my scholarly obligations, I would like to go over my reasons.\\

Firstly, I don't respect the methods of formal linguistics. The extent of my prior exposure to formal linguistics is reading Heim \& Kratzer. I have a copy signed by Angelika, with an inscription "I hope it wasn't pure torture." It was unreadable because rather good and simple ideas were obfuscated in the mathematical equivalent of machine code. The first half of the book up to quantifiers can be summarised as a cartesian closed functor [Masters], and here is a fresh and morally lossless categorical exegesis of the technical contents of the whole book: Define \emph{"generative grammar"} to be a coloured operad, define semantic data to be a coloured operad along with an algebra valued in the homsets of cartesian closed category generated by types $e$ and $t$, and define \emph{"semantics for generative grammar"} to be the obvious operad map. I am even being generous, by already generalising in that description beyond truth-conditional semantics for natural language, which have been untenable since Putnam's permutation argument was fixed by Barr [Putnam, Barr]. Since I had it on good authority that Heim \& Kratzer was a gold standard, I made the informed decision to engage with all formal linguistics as little as possible to avoid mental pollution. I believe that I have adequately illustrated the possibilities of using diagrammatic mathematics in formal linguistics, and that my disdain for the extant formal alternatives is valid.\\

Secondly, I think the field as a whole is doomed anyway, at least spiritually. I think it is a deep kind of irony that the field that nominally concerns itself with meanings is suffering from meaninglessness, and I'm fairly certain it isn't just projection on my part. Insofar as:
\begin{enumerate}
\item large language models are \emph{the} hard technology for language
\item hard sciences are so by virtue of hard technologies
\item formal linguistics was destined to be a hard science
\item large language models have nothing to do with formal linguistics
\end{enumerate}
The whole thing has been a depressing failure of existential proportions. As far as I can tell, nobody contests point 4; in fact, the linguists I have spoken with have taken pains to stress this point [????]. Point 3 is just too embarrassing to admit out loud, but it is difficult to articulate why that is so without the whole complex. Point 2 is an unspoken and fairly modern cultural convention; for the vast majority of human history, craft has been ahead of science [Architecture]. Hence, the strategy is to deny point 1, which is why any self-respecting computational linguist will hasten to point out a small handful of pet examples where LLMs fail. I am even willing to grant that the failures are real and systematic and peer-reviewed; there is just an air of defensiveness and commiseration that is difficult to look past, especially in contrast with the vital exuberance on the other side.

So what went wrong? As soon as mathematics became involved, the aim became procedural knowledge of language, and the grail was a language computer. It is impossible to take any other stance seriously: who would dare admit that the true aim was to establish an eternal monastic tradition of artisinal Markerese? As I have also indicated in the introduction, the relationship of theory to empirical capture is complicated today, because while the gradual accommodation of empirical observation is how a theory grows, the observed conclusion-at-scale of star-unstar is a theory-free big-data situation [BigData].

The only value system in which I see the endeavour of formalising the structure of language as a worthwhile activity is some kind of authentic postironic neoromanticism that promotes a sort of aesthetic, productive, and communicative diversity of formalisms and methods of inquiry rather than canonical or universal structures, but on this view I take the vast majority of activity within any given scholarly institution as seriously as, say, Mario 64 speedrunning [A-button].

\subsection{Sounds good, doesn't work}

(synthesis via diagrams; open questions: how to save both logical structure and vectoriality(context/higher-order)?)