\section{A Partial History of String Diagrams}

\newthought{Observation 1: We have a primitive urge to depict.}

\[cavedrawings to language\]

\newthought{Observation 2: Natural language presents itself as a universal interface for thought.}

\[placeholder\]

\newthought{The pictorial capture of language and thought is a cursed dream.} We suffer a seductive alchemical urge to capture the essence, laws and structure of things in words and pictures -- containment in symbol and form. On occasion, this urge is focused on itself as an ouroboros; a doomed endeavour to use the faculties of abstraction, representation, and structure to consume itself.

\[peirce\]

\[magnusson\]

\[maybe jung\]

\newthought{It is our turn to suffer this sickness.} The difference this time around is in approach: we use \emph{string diagrams}. If depiction is the domain of geometry -- the mathematics of form in space, and if language is the domain of algebra -- the mathematics of symbolic manipulation, then string diagrams are our bridge. We recall a quote by the late Michael Atiyah:\\

\emph{Algebra is the offer made by the devil to the mathematician. The devil says: "I will give you this powerful machine, it will answer any question you like. All you need to do is give me your soul: give up geometry and you will have this marvellous machine."}\\

String diagrams are a loophole in the devil's contract. Descartes' led an incursion into geometry with algebra, and string diagrams are the belated counteroffensive, or peaceful synthesis.

\newthought{String diagrams are formal representation and reasoning systems for monoidal categories.} "Formal" distributes over "representation" and "reasoning", which arose chronologically in that order.

\subsection{Formal visual representation}

Pre-formal diagrams are about visual expression. By the restrictions of writing technologies, the kind of visual representations we are interested in are two-dimensional. So visual means geometric in the plane. The ancient ancestors of modern string diagrams are systems for visual representation of formal symbolic constructions. For these ancestral diagrams, reasoning, if there is any, takes place elsewhere in the symbolic realm, not within and between diagrams.\\

\[Euclid: Geometry\]

\[Venn \& Carroll: Syllogisms\]

\[Petri: Chemistry\]

\subsection{Convergent Evolution}

String diagrams arise naturally when keeping track of many pairwise connections between entities becomes unwieldy for one-dimensional syntax. The usual evolutionary path is to first adopt Einstein notation with indexed superscripts and subscripts for bookkeeping, followed by the depiction of connected indices as wires in the plane.

\newthought{In Physics}

\newthought{In Logic}

\newthought{In Computer Science}

\subsection{Formal visual reasoning}

